Constraining gravitational wave amplitude birefringence with GWTC-3

### 2023-04-17 src/tex/ms.tex
Minor typo fixes, additional reference for re-weighting posteriors.

By implementing Eq.~\eqref{eq:waveform_modification}, which is a first-order approximation of the birefringence model, we generally break the degeneracy between birefringence and source orientation; this was also the case in the frequency-dependent relations studied in \cite{Yamada_2020,Wang_2021}.
Nevertheless, there exist systems for which the degeneracy cannot be broken in practice because not enough frequencies are available in the data.
This may be the case for quasimonochromatic sources, like nonaxisymmetric pulsars or very light binaries, which are well approximated by a single Fourier mode.

@@

We estimate source parameters using a custom version of the \textsc{Bilby} software \citep{Bilby}, modified from the baseline version to apply Eq.~\eqref{eq:waveform_modification} for any \ac{GR} baseline waveform.
We take the \ac{PE} configuration in \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset} as a starting point, with \textsc{IMRPhenomXPHM} \citep{Pratten:2020ceb} as the reference waveform.
We apply a distance prior corresponding to a uniform distribution over comoving volume and source-frame time, and set the prior on $\kappa$ to be uniform between $-1$ and $1$.

@@

As a visual check for outliers, we reconsider the set of measurements in Fig.~\ref{fig:violin_kappa} in light of the hierarchical result for $\mu$ and $\sigma$ in Fig.~\ref{fig:corner_Gaussian}, including GW200129 (blue curve).
This amounts to reweighting the $\kappa$ posterior for each event under a population prior marginalized over $\mu$ and $\sigma$, conditional on the measurements from all other events \cite{Callister:T2100301}.
The result in Figure \ref{fig:reweighted_kappa} does not show evidence for any of the events being in obvious tension with the population, even though the GW170818 curve stands out from the rest due to its higher support for $\kappa > 0$.
This feature appears to offset a few other events which tend to favor $\kappa < 0$.
The interaction between these distributions leads to a hyperposterior that is fully consistent with $\mu = 0$ while offering some support for $\sigma > 0$ (Fig.~\ref{fig:corner_Gaussian}).
Future observations will determine whether there is truly evidence for a nonvanishing variance in this population.

>>>>

By implementing Eq.~\eqref{eq:waveform_modification}, which is a first-order approximation of the birefringence model, we generally break the degeneracy between birefringence and source orientation; this was also the case in the frequency-dependent relations studied in \cite{Yamada_2020,Wang_2021}.
Nevertheless, there exist systems for which the degeneracy cannot be broken in practice because not enough frequencies are available in the data.
This may also be the case for quasimonochromatic sources, like nonaxisymmetric pulsars or very light binaries, which are well approximated by a single Fourier mode.

@@

We estimate source parameters using a custom version of the \textsc{Bilby} software \citep{Bilby}, modified from the baseline version to apply Eq.~\eqref{eq:waveform_modification} for any \ac{GR} baseline waveform.
We take the \ac{PE} configuration in \citep{GWTC-2.1, GWTC-3, GWTC-2.1_dataset, GWTC-3_dataset} as a starting point, with \textsc{IMRPhenomXPHM} \citep{Pratten:2020ceb} as the reference waveform.
We apply a distance prior corresponding to a uniform distribution over comoving volume and source-frame time, and set the prior on $\kappa$ to be uniform between $-1 \, \mathrm{Gpc}^{-1}$ and $1 \, \mathrm{Gpc}^{-1}$.

@@

As a visual check for outliers, we reconsider the set of measurements in Fig.~\ref{fig:violin_kappa} in light of the hierarchical result for $\mu$ and $\sigma$ in Fig.~\ref{fig:corner_Gaussian}, including GW200129 (blue curve).
This amounts to reweighting the $\kappa$ posterior for each event under a population prior marginalized over $\mu$ and $\sigma$, conditional on the measurements from all other events \cite{Miller2020,Callister:T2100301}.
The result in Figure \ref{fig:reweighted_kappa} does not show evidence for any of the events being in obvious tension with the population, even though the GW170818 curve stands out from the rest due to its higher support for $\kappa > 0$.
This feature appears to offset a few other events which tend to favor $\kappa < 0$.
The interaction between these distributions leads to a hyperposterior that is fully consistent with $\mu = 0$ while offering some support for $\sigma > 0$ (Fig.~\ref{fig:corner_Gaussian}).
Future observations will determine whether there is truly evidence for a nonvanishing variance in this population.
